### 2016-09-20 doc/manuscript.Rmd
Detailed PLS results more (incl BCF) + edit methods.

- Run PLS with just those FA that loaded strongly to see the predictive ability.

@@

- Primary analysis:
    - Look over medications, how they influence the TAGFA
    - Heatmap
    - GEE
        - SA for the large outlier for the TAGFA (see dist plot)
        - Include # of both uncorrected and corrected (FDR) significant associations
        - QIC modeling
        - DAG development
        - Look over the entire model (see which variables are significant, etc)
        - Sensitivity analysis with Carb intake? And with adiponectin? (which is
        involved in TAG metabolism)
    - PLS with ISI and HOMA-IR
        - CV for determining number of comps (best is 2)
        - Loadings (of TAGFA) (tho all are the same, use HOMA/ISI. BCF have
        lower corr between measured and predicted) (Use COrr of both components
        (1 and 2))
        - Prediction (of predicted vs real) (in the pls.rmd file, not in
        manuscript (or maybe in ESM?))
        - Use extracted PLS components in GEE (?)
        - Train and test (CV) to see how it predicts in a new dataset (ie. a
        future visit) (done. In pls.rmd file. As ESM?)
- Secondary analysis:
    - PLS-DA (dysgly)
        - Train and test (CV)
        - Loadings (of TAGFA) 
        - Groupings
        - Prediction (misclassification)
    - AROC
        - dysgly with clinical TAG vs the 'best' fatty acid to compare to

@@

# Correlation between predicted and actual
pls_homa_t <- analyze_pls(y = 'linvHOMA')
pls_isi_t <- analyze_pls(y = 'lISI')

@@

Insulin sensitivity and beta-cell function indices were computed using the OGTT 
glucose and insulin data. Insulin sensitivity was assessed using 1/HOMA-IR (1
divided by HOMA-IR) [@Matthews1985] and the Insulin Sensitivity Index (ISI)
[@Matsuda1999]. HOMA-IR largely reflects hepatic insulin resistance, while ISI
reflects whole-body insulin sensitivity [@AbdulGhani2007]. Beta-cell function
was assessed using the Insulinogenic Index [@Wareham1995] over HOMA-IR (IGI/IR)
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2) [@Retnakaran2009]. IGI/IR
is a measure of the first phase of insulin secretion while ISSI-2 is analogous to
the disposition index (but using OGTT values). Each index has been validated
against gold standard measures [@Matthews1985; @Matsuda1999; @Retnakaran2009].

The primary outcome variables for this analysis were 
1/HOMA-IR, ISI, IGI/IR, and ISSI-2, which were log-transformed for
the statistical modeling. The primary predictor variables for this analysis were
22 individual TAGFA as mole percent (mol%) of the total fraction and as
concentration (nmol/mL). Pearson correlation coefficients were computed to
assess the relationships of individual TAGFA with other continuous variables.

Of note is clinically measured TAG was only quantified on a portion of the
6-year visit (n = ~120 {{ find true number }}).

@@

tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
female <- calc_discr_npct(ds0yr$Sex, 'Female')
euro <- calc_discr_npct(ds0yr$Ethnicity, 'European')
fhd <- calc_discr_npct(ds0yr$FamHistDiab, 'Yes')
est0 <- calc_gee_estci(gee_unadj_df)
est1 <- calc_gee_estci(gee_adj_df)
fa_cor <- calc_cor()
cor_pls_homa <- calc_pred_corr(pls_homa_t, pls_homa_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
```

@@

Clustering analysis using PLS, shown in `r fig_pls`, revealed that 14:0, 14:1n-7,
16:0, and 16:1n-7 loaded strongly (meaning the FA fell between the dashed and
solid circles) and negatively on 1/HOMA-IR and ISI, suggesting this cluster of
TAGFA tracks together with lower insulin sensitivity. The TAGFA 20:2n-6,
20:5n-3, 22:5n-3, and 22:6n-3 loaded positively on both insulin sensitivity 
measures. No other TAGFA loaded strongly, and the results of which remained
after running the PLS model on the testing set. The predictive ability of the 
PLS model, determined by correlating the predicted outcome values with the actual 
values, were good (1/HOMA-IR: `r cor_pls_homa$r_p`; ISI: `r cor_pls_isi$r_p`).

@@

@@

>>>>

- Run PLS with just those FA that loaded strongly to see the predictive ability.
- SA with/without NEFA
- SA with/without WC

@@

- Primary analysis:
    - Look over medications, how they influence the TAGFA
    - GEE
        - SA for the large outlier for the TAGFA (see dist plot)
        - Include # of both uncorrected and corrected (FDR) significant associations
        - QIC modeling
        - DAG development
        - Look over the entire model (see which variables are significant, etc)
        - Sensitivity analysis with Carb intake? And with adiponectin? (which is
        involved in TAG metabolism)
    - PLS with ISI and HOMA-IR
        - Use extracted PLS components in GEE (?)
- Secondary analysis:
    - AROC
        - dysgly with clinical TAG vs the 'best' fatty acid to compare to

@@

# Correlation between predicted and actual
pls_homa_t <- analyze_pls(y = 'linvHOMA')
pls_isi_t <- analyze_pls(y = 'lISI')
pls_igi_t <- analyze_pls(y = 'lIGIIR')
pls_issi2_t <- analyze_pls(y = 'lISSI2')

@@

Insulin sensitivity and beta-cell function indices were computed using the OGTT 
glucose and insulin data. Insulin sensitivity was assessed using 1/HOMA-IR (1
divided by HOMA-IR) [@Matthews1985] and the Insulin Sensitivity Index (ISI)
[@Matsuda1999]. HOMA-IR largely reflects hepatic insulin resistance, while ISI
reflects whole-body insulin sensitivity [@AbdulGhani2007]. Beta-cell function
was assessed using the Insulinogenic Index [@Wareham1995] over HOMA-IR (IGI/IR)
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2) [@Retnakaran2009]. IGI/IR
is a measure of the first phase of insulin secretion while ISSI-2 is analogous to
the disposition index (but using OGTT values). Each index has been validated
against gold standard measures [@Matthews1985; @Matsuda1999; @Retnakaran2009].
Specific formulas for each OGTT-derived measure can be found in the {{
Supplementary Material }}.

The primary outcome variables for this analysis were 1/HOMA-IR, ISI, IGI/IR, and
ISSI-2, which were log-transformed for the statistical modeling. The primary
predictor variables for this analysis were 22 individual TAGFA as mole percent
(mol%) of the total fraction and as concentration (nmol/mL). Clinically-measured
TAG is also included as a primary predictor to serve as a comparison against the
TAGFA composition. A limitation with the clinically-measured TAG is that it was
only quantified on a portion of the 6-year visit (n = ~120 {{ find true number
}}). Pearson correlation coefficients were computed to assess the relationships
of individual TAGFA with other continuous variables.

@@

tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
female <- calc_discr_npct(ds0yr$Sex, 'Female')
euro <- calc_discr_npct(ds0yr$Ethnicity, 'European')
fhd <- calc_discr_npct(ds0yr$FamHistDiab, 'Yes')
est0 <- calc_gee_estci(gee_unadj_df)
est1 <- calc_gee_estci(gee_adj_df)
fa_cor <- calc_cor()
cor_pls_homa <- calc_pred_corr(pls_homa_t, pls_homa_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
cor_pls_igiir <- calc_pred_corr(pls_igiir_t, pls_igiir_t$test_data)
cor_pls_issi2 <- calc_pred_corr(pls_issi2_t, pls_issi2_t$test_data)
```

@@

To determine which components to use, internal CV showed that the first two
components gave the highest amount of explained variance (data not shown).
Confirming the findings from the GEE models, PLS analysis on the beta-cell
function measures revealed that there was a low correlation (though significant
at p`r cor_pls_igi$p`) between the predicted values and the actual values (r=
`r cor_pls_igi$r` to `r cor_pls_issi2$r`), suggesting poor predictability. While 
the clustering of TAGFA from the PLS results of the beta-cell function measures 
were similar to the insulin sensitivity measures, only the insulin sensitivity 
measures are presented given the low predictability.

The PLS results for the insulin sensitivity (`r fig_pls`) revealed a clustering 
of the FA 14:0, 14:1n-7, 16:0, and 16:1n-7. These TAGFA loaded strongly (meaning
the FA fell between the dashed and solid circles) and negatively on 1/HOMA-IR
and ISI, suggesting this cluster of TAGFA tracks together with lower insulin
sensitivity. The TAGFA 20:2n-6, 20:5n-3, 22:5n-3, and 22:6n-3 loaded positively
on both insulin sensitivity measures. No other TAGFA loaded strongly, the
results of which remained after running the PLS model on the cross-validated
testing set. The predictive ability of the PLS model, determined by correlating
the predicted outcome values with the actual values, were good (1/HOMA-IR: 
`r cor_pls_homa$r_p`; ISI: `r cor_pls_isi$r_p`).

@@

[@Barrows2006a] NEFA contributes 60-80% of newly synth VLDL-TAG during fasting
and ~40% during fed. Dietary FA contribute ~25% to newly synth VLDL-TAG.
VLDL-TAG is not cleared (lipolyzed) as much as chylomicron and so stays in
circulation longer (hence why there is bigger TAG pool vs NEFA)

@@

Include DAGitty image
